BACI of biomass change for VOCC source cells

### 2019-08-22 analysis/VOCC/4-vocc-BACI.Rmd
exact matching option

```{r}
vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )

@@

#browser()
  alldata <- rbind(source_biomass, control_biomass) 
  #alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0))
  #fit <- McSpatial::matchprop(cell_type ~ start_density + depth, data= alldata, ytreat = "source")
  fit_formula <- paste0("cell_type ~ start_density + depth") 
 # fit_formula <- paste0("cell_type ~ start_density + depth + trawled + muddy + any_rock + omega_s")

  for (j in (variable_names)) {
  fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  }

  matched <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = "source")

  matched_time0 <- matched %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- matched %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)

@@

Prep data
```{r}
data <- vocc_by_sp[["adult"]]
#View(data)

data$log_density <-  log(data$density)
data$distance2 <- data$distance^2

@@

Plot data
```{r}
lapply(variable_names, function(j) {
ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") 
})

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)
```

Basic BACI
```{r}
for (j in (variable_names)) {
  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + distance + distance2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
}

@@

```{r}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,7,5,6,3,4), title = paste(species, "adult"))
```

@@

for (j in (variable_names)) {
  gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
}

gam4 <- gamm4::gamm4(as.formula(gamm_formula),random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

@@

summary(gam4$gam) ## summary of gam
summary(gam4$mer) ## underlying mixed model
anova(gam4$gam)

@@

plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density") 
```

@@

Prep data
```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)
idata$distance2 <- idata$distance^2

@@

Plot data
```{r}
lapply(variable_names, function(j) {
ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") 
})
ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

@@

GAMM
```{r}
# for (j in (variable_names)) {
#   gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
# }

@@

plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density") 
```

Misc exploring gamm 
```{r eval=FALSE}
# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
# summary(mi)
# 
# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
# summary(mi)
# mi
# 
# 
# idata$after <- as.factor(idata$after)
# idata$matchobs <- as.factor(idata$matchobs)
# 
# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(distance) + #s(X) + s(Y) + 
#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
# 
# 
# plot(m4b$gam,pages=1)
# 
# summary(m4b$gam) ## summary of gam
# summary(m4b$mer) ## underlying mixed model
# anova(m4b$gam) 
# 
# gamtabs(m1)
# 
# #library(itsadug)
# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
# 
# library(mgcv)
```

>>>>

```{r}
#i <- biomass_list[[1]]

vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )

@@

#browser()
  alldata <- rbind(source_biomass, control_biomass) 
  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
    select(-var_2_min, -var_2_max) %>% 
    rename(vect_dist = distance)
  alldata$treat <- as.factor(alldata$treat)
  ggplot(alldata, aes(x=cell_type, y=log(start_density))) + geom_boxplot()

  ## still including this trimming to reduce size of dataset 
  control_min <- min(alldata$start_density[alldata$treat=="0"])
  source_min <- min(alldata$start_density[alldata$treat=="1"])
  # control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
  # source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
  group_min <- max(control_min, source_min)

  control_max <- max(alldata$start_density[alldata$treat=="0"])
  source_max <- max(alldata$start_density[alldata$treat=="1"])
  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 0.95)
  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 0.95)
  group_max <- min(control_max, source_max)
  # 
   alldata <- filter(alldata, start_density >= group_min) #%>% filter(start_density <= group_max)
  # 
  # num_source <- nrow(alldata[alldata$treat=="1", ])
  #

  # alldata <- filter(alldata, start_density >= control_min)

  # low_biomass <- quantile(alldata$start_density, 0.1)
  # alldata <- filter(alldata, start_density >= low_biomass )

  # if (order == "decreasing") 
  #alldata <-   alldata[rev(order(alldata$start_density)), ]
  # 
  # if (order == "increasing") 
 # alldata <- alldata[rev(order(alldata$start_density)), ]

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
  alldata$depth.st <- arm::rescale(log(alldata$depth)) 
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock) 
  #alldata$omega_s.st <- arm::rescale(alldata$omega_s)

 ## Using mysterious cem optimization ... doesn't work   
  # fit_formula <- paste0("treat ~ start_density_st") 
  # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
  # fit_formula <- paste0("treat ~ start_density_st + depth + trawled + muddy + any_rock + omega_s + x + y") 
  # fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") #+ + omega_s
  # 
  # # for (j in (variable_names)) {
  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  # #   }
  # 
  # # # options("optmatch_max_problem_size" = Inf)
  # # matched <- MatchIt::matchit(as.formula(fit_formula), data = alldata, method = "optimal", ratio = 1)
  # #fit_formula <- paste0("treat ~ depth.st")
  # matched <- MatchIt::matchit(as.formula(fit_formula),
  #   method = "cem",
  #   distance = "probit",
  #   discard="control",
  #   data = alldata)

  # Using exact match and rounding
  fit_formula <- paste0("treat ~ depth.st +  muddy.st + any_rock.st") #+ trawled + omega_s
  for (j in (variable_names)) {
    fit_formula <- paste0("", fit_formula, " + ", j, "_s")
    }

  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st"),
    method = "nearest", m.order = "random",
    #method = "full",
    distance = "probit",
    discard="control",
    data = alldata)

  mdata <- MatchIt::match.data(matched, distance = "pscore")

  summary(matched)
  plot(matched)

  sources <- mdata[row.names(matched$match.matrix), ]
  sources$origobs <- row.names(sources)   
  sources$matchobs <- matched$match.matrix[,1]

  controls <- mdata[matched$match.matrix, ]
  controls$origobs <- row.names(controls)   
  controls$matchobs <- matched$match.matrix[,1]

  # controls2 <- mdata[matched$match.matrix, ]
  # controls2$origobs <- row.names(controls)   
  # controls2$matchobs <- matched$match.matrix[,2]

  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")

  #mdata <- mdata %>% rename(distance = vect_dist) #%>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)

  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)

@@

Prep data
```{r}
data <- vocc_by_sp [["adult"]]
# View(data)

data$log_density <-  log(data$density)
data$vect_dist2 <- data$vect_dist^2

@@

Plot data
```{r}
ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
  geom_point(alpha=0.4, size =4) +
  scale_shape_manual(values= c(4, 1)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  theme_bw()

lapply(variable_names, function(j) {
ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  theme_bw()
})

ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  facet_wrap(~time) +
  theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()
```

Basic BACI
```{r}
for (j in (variable_names)) {
  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
}

@@

```{r}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
```

@@

for (j in (variable_names)) {
  gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
}

gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

# for (j in (variable_names)) {
#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + te(X,Y)")
# }
# gam <- mgcv::gamm(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

@@

summary(gam4$mer) ## underlying mixed model
anova(gam4$gam)

@@

plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density")

summary(gam4$gam) ## summary of gam
```

@@

Prep data
```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)
idata$vect_dist2 <- idata$vect_dist^2

@@

Plot data
```{r}
ggplot(filter(data, as.numeric(matchobs) <= 150), aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
  geom_point(alpha=0.4, size =4) +
  scale_shape_manual(values= c(4, 1)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  theme_bw()

lapply(variable_names, function(j) {
ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") 
})
ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

@@

GAMM
```{r}
# for (j in (variable_names)) {
#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
# }

@@

#my_cols <- c("#E69F00", "#56B4E9")
plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density"
  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
  ) 
```

Misc exploring gamm 
```{r eval=FALSE}
# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
# summary(mi)
# 
# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(vect_dist) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
# summary(mi)
# mi
# 
# 
# idata$after <- as.factor(idata$after)
# idata$matchobs <- as.factor(idata$matchobs)
# 
# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(vect_dist) + #s(X) + s(Y) + 
#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
# 
# 
# plot(m4b$gam,pages=1)
# 
# summary(m4b$gam) ## summary of gam
# summary(m4b$mer) ## underlying mixed model
# anova(m4b$gam) 
# 
# gamtabs(m1)
# 
# #library(itsadug)
# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
# 
# library(mgcv)
```
